TBA

### 2019-09-26 paper.tex
Restarting paper, added data processing

\usepackage{booktabs}
\usepackage[ruled]{algorithm2e}

\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage[caption=false]{subfig}
\usepackage[capitalise]{cleveref}
\usepackage{todonotes}

@@

\title{TBA}

\author{TBA}

%\title{Bone-Conduction Audio Interface to Guide People with Visual Impairments}

%\author{Jacobus C. Lock\inst{1} \and 
%Iain D. Gilchrist\inst{2} \and
%Grzegorz Cielniak\inst{1} \and
%Nicola Bellotto\inst{1}}

%\institute{University of Lincoln, Lincoln, LN6 7TS, UK \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk} \and
	   %University of Bristol, Bristol, BS8 1TH, UK \email{i.d.gilchrist@bristol.ac.uk}}

%\authorrunning{Jacobus C. Lock et.\ al.}
%\titlerunning{Bone-conduction Interface For People with Visual Impairments}

\maketitle
%\setcounter{footnote}{0}

\begin{abstract}
  The ActiVis project's aim is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
  This system uses bone-conduction headphones to transmit audio signals to the user and requires an effective non-visual interface.
  To this end, we propose a new audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
  The vertical position on the median plan is given by adjusting the tone's pitch to overcome the audio localisation limitations of bone-conduction headphones. 
  This interface is validated through a set of experiments with blindfolded and visually impaired participants. 
%\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch \and bone-conduction}
\end{abstract}

@@

In recent years, governments have spearheaded numerous initiatives to support people with disabilities and enable them to play a more active role in modern society.
The UK's Royal National Institute of Blind People (RNIB) for example has prioritised improving access to everyday services and products, such as public transport and mobile apps~\cite{rnib2016uk}.
Improvements in modern computing have made it possible for new and innovative solutions to these problems to come to the fore.
In particular, researchers in the active vision field have have made much progress in enabling machines to autonomously manipulate cameras to gather information about an environment for mapping and object finding tasks~\cite{bajcsy2018revisiting,lock2019active}.
There is, however, a significant research question about whether techniques from active machine vision can be applied to humans, i.e.\ can a machine identify a point of interest in a scene and direct a human, instead of an electronic servo, to focus on that point?
If this can be done, it would be beneficial to people with visual impairments and will augment their ability to search for an arbitrary point or object of interest and identify an unknown scene.

The ActiVis\footnote{https://lcas.github.io/ActiVis/} project aims to deliver a mobile guidance system that will ultimately be able to guide a user with vision impairments on the last leg of their journey, i.e.\ the so-called `last 10-yard problem'. 
The prototype platform is based on a Google Project Tango\footnote{https://en.wikipedia.org/wiki/Tango\_(platform)} device, pictured in \cref{fig:tango} and \cref{fig:tango-headphone}, that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities. 
It also provides access to Android's full range of interface tools and I/O options. 
Furthermore, a set of bone-conduction headphones (pictured in \cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.

\begin{figure}[t]
  \centering
  \subfloat[]{\label{fig:tango}\includegraphics[width=0.4\columnwidth]{figures/tango_asus.jpeg}}
~
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.3\columnwidth]{figures/tango_headphone.png}}
~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.2\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the latest Tango mobile phone (left), the former Tango tablet and the bone-conduction headphones used in this work (centre), and a participant with visual impairments during an experiment (right).}
\end{figure}

Humans are able to determine the 3D position of a sound source and by exploiting this natural ability, the real-time guidance instructions can easily be interpreted without posing a significant cognitive load.
A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
A similar approach was used in~\cite{durette2008visuo}, but the study did not focus on or investigate the interface's efficacy.
Indeed, the work presented in this paper serves as an initial study into an interface that can effectively communicate the pan and elevation angles of a target location through a set of bone-conduction headphones using standard audio libraries and audio manipulations.
Furthermore, we demonstrate this approach's effectiveness through a simple experiment.

The rest of the paper starts by discussing previous relevant works and research in \cref{sec:prev-work}, followed by a discussion on the design and implementation of our interface in \cref{sec:interface-design}.
This is followed by a description of the experiments that were conducted and a discussion of their results in \cref{sec:experiment-and-results}.
Finally, the paper concludes with a summary and a short discussion on future research prospects in \cref{sec:conclusion}.

\section{Previous Work}\label{sec:prev-work}

Multiple mobile navigation and travel aids have been devised throughout the years, some as part of a commercial undertaking and many as part of academic research.
The majority of these systems rely on one or a combination of vocal~\cite{mocanu2016when,chessa2016integrated,kanwal2015navigation}, audio~\cite{schwarze2015intuitive,rodriguez2012obstacle,katz2010navig} and haptic~\cite{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to communicate guidance instructions to the user, each with their own sets of features and limitations.
In general, participants with visual impairments report that they prefer haptic and vocal feedback out of the three options~\cite{arditi2013user}.
However, haptic feedback systems typically have extra hardware requirements to transmit the guidance instructions.
Furthermore, haptic and vocal feedback can become a cognitive burden, particularly where high resolution guidance, which can exceed the bandwidth of these particular senses, is required.
Indeed, participants report that they would prefer control over the vocal feedback channel and trigger guidance instructions, instead of being given constant guidance instructions~\cite{arditi2013user}.
Simple audio tones are less affected by these bandwidth and hardware limitations, but they can potentially fatigue the user's senses if too unpleasant.

Researchers have investigated using audio signals that are spatialised with a head-related transfer function (HRTF), simulating a sound source located at some arbitrary 3D position~\cite{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
The authors generally report favourable results for this guidance approach when used with standard over-ear headphones or speakers. 
However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}. 
This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\cite{stanley2006lateralization}.
In~\cite{durette2008visuo}, they transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
However, these studies focussed on their system's total effectiveness, rather than the effectiveness of the pitch adjustment mechanism alone. In the next sections we want to investigate this more closely.

\section{Interface Description}\label{sec:interface-design}

\subsection{Hardware Selection}

Electronic navigation aids and guidance systems for people with vision impairments often have acceptability and usability problems due to limitations in academic research and prototypes, such as prohibitive costs, unfriendly user interfaces and significant hardware requirements (e.g.\ head-mounted cameras and GPS antennae)~\cite{golledge2004stated,yusif2016older,arditi2013user}.
Indeed, the traditional walking cane remains the standard tool for navigation and obstacle avoidance. 
With the ActiVis system, we aim to tackle the issue of cost, user-acceptance and usability by using only off-the-shelf devices and hardware, as far as possible, and avoiding any special requirements.
Following this design principle, we adopted a hand-held mobile device and a discreet headset to avoid the person standing out in public places while using the system.
The ActiVis system is based on a concept proposed in~\cite{lock2017portable,lock2019active} that uses a Google Tango device that is able to localise itself in real-time and has the benefit of a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.

@@

\subsection{Human Audio Localisation}

@@

\subsection{Interface Design}

@@

\subsubsection{Pan}

@@

\subsubsection{Elevation}

@@

@@

\section{Experiments}\label{sec:experiment}

To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
The participants are given a Tango device running an app written for the experiment that generates a set of virtual targets and presents them to each participants, one at a time. 
The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
When a participant was confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localisation API.\
A total of 28 targets were generated per participant.

Two groups of participants were recruited for the experiments. 
Group \textit{G1} consisted of 10 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 4 people with severe visual impairments. 
Both groups were given some time before the experiments to familiarise with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets, without worrying about the time it took.

\section{Results}\label{sec:results}

The data collected during the experiment were clustered together by participant groups and analysed.
Similar to the results from~\cite{macdonald2006spatial,schonstein2008comparison}, we use the absolute mean errors for each dimension, which allows us to compare our results to the aforementioned works'. 
These results are presented in the boxplots in \cref{fig:err-results-fig} and summarised in \cref{tab:results}.

\begin{figure}
  \centering
  \subfloat[]{\label{fig:err-results-abs}\includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err_abs.png}}
\quad
  \subfloat[]{\label{fig:err-results}\includegraphics[width=0.7\columnwidth]{figures/pan_tilt_err.png}}
  \caption{Boxplots of the absolute (top) and signed (bottom) angular errors collected during the pointing experiment. }\label{fig:err-results-fig}
\end{figure}

\begin{table}
  \centering
  \caption{A summary of the results collected from the experiment. }\label{tab:results}
  \begin{tabular}{p{1cm}p{2cm}p{2.2cm}@{\hskip 1cm}p{3.2cm}}
    \toprule
    &           & Mean absolute error [rad]  & Spearman Correlation score \\ \midrule
    \multirow{2}{*}{\textit{G1}} & Pan       & $0.23\pm0.24$ & $0.88\;(p<0.001)$ \\
				 & Elevation & $0.25\pm0.28$ & $0.64\;(p<0.001)$ \\ \midrule
    \multirow{2}{*}{\textit{G2}} & Pan       & $0.45\pm0.21$ & $0.41\;(p<0.001)$ \\
				 & Elevation & $0.36\pm0.32$ & $0.61\;(p<0.001)$ \\
    \bottomrule
  \end{tabular}
\end{table}

The average pan error from group \textit{G1} falls well within the ranges observed in~\cite{macdonald2006spatial,schonstein2008comparison} of $[0.16, 0.38]$ radians and $[0.17, 0.26]$ radians respectively. 
Indeed, the majority of samples fall even within the latter, more conservative range. 
However, the participants with visual impairments in group \textit{G2}  demonstrate a wider spread in error data and higher average error than \textit{G1} (0.23 vs. 0.45 radians, Kruskal-Wallis test $p<0.001$). 
The results in~\cite{katz2011spatial,zwiers2001spatial} reported a similar trend, although an investigation with a larger sample size for \textit{G2} should be considered before definitive conclusions can be made.

The elevation estimation performance for both groups deteriorated when compared to their pan estimation results, which was somewhat expected given previous experimental results on human audition~\cite{barfield1997visual}. 
The mean absolute errors of the participants are \SI{0.25}{\radian} and \SI{0.36}{\radian} for groups \textit{G1} and \textit{G2}, respectively.
As with the pan estimation errors, the participants from \textit{G1} seemed more adept at estimating the correct elevation.
Comparing these results to methods that convey elevation through an HRTF, we see an improvement of approximately 63\% and 47\% for each group~\cite{schonstein2008comparison}. 
Indeed, the performance is comparable to that of open-back and expensive in-ear headphones.
There is a more modest improvement of 43\% in \textit{G1} and 16\% for \textit{G2} when compared to the individualised and adjusted HRTFs in~\cite{stanley2006lateralization}.
However, all the participants in~\cite{stanley2006lateralization} had healthy eyesight, making the latter comparison invalid. 
Interestingly, the elevation errors for both groups are centred around the negative angle, as shown in \cref{fig:err-results}, indicating a possible underlying bias either in the participants or the interface.
A similar trend was observed in~\cite{stanley2006lateralization}. Further investigation with a larger participant pool is needed to determine this underlying bias, if any.

\begin{figure}
  \centering
  \includegraphics[width=1.0\columnwidth]{figures/correlation.png}
  \caption{Scatter plots that show the targets' actual positions and the participants' guesses, as well as the line of best fit showing the relationship between the data. }\label{fig:correlation-results}
\end{figure}

The error data was used to generate scatter plots that show the correlation between the target's direction presented by the guidance system to the participants and the one they thought the targets were.
These plots are given in \cref{fig:correlation-results} and show a reasonably positive correlation between the guidance and the selected targets, indicating that the participants correctly interpreted the interface's output in most of the cases. 
The Spearman correlation scores for each dataset, listed in \cref{tab:results}, confirm this, as well as the negative angle bias in the elevation dimensions.

\section{Conclusion}\label{sec:conclusion}

In this work, we proposed a new audio interface that conveys a target's pan and elevation angles through a spatialised audio tone.
The tone's pitch is adjusted to convey the elevation and compensate for the shortcomings of bone-conduction headsets in transmitting the elevation component of a spatialised signal.
The results obtained from a simple pointing experiment are encouraging as they show improved performance over purely spatialised signals transmitted via bone-conduction headphones.
Furthermore, the limitations of the latter were largely overcome by communicating the target's elevation with the tone's pitch, achieving pan and elevation errors comparable to that of normal over-ear headphones.

Future work should consider a more in-depth investigation with a larger participant pool and a more detailed statistical analysis.
Interesting questions include whether adjusting the gradient of the elevation function would make it easier for the user to guess the actual target's elevation, and to what degree the performance of participants with limited and healthy eyesight differs.
Furthermore, experiments with a large group of people with visual impairments would provide an opportunity to gauge their opinion on the interface and identify possible avenues of improvements.
Finally, it would be interesting to incorporate a co-adaptation mechanism between human and assistive device, as suggested in~\cite{gallina2015progressive}, to improve its usability and performance over time.

\paragraph{{\bf Acknowledgements.}}\label{sec:acknowledge}
This research is partly supported by a Google Faculty Research Award.
We would like to thank the Voluntary Centre Services UK for their help in facilitating the experiments with people with limited vision.

\bibliographystyle{splncs04}
\bibliography{bib}

>>>>

@@

@@

\section{Previous Work}

\section{System Description}

Existing electronic navigation aids have typically struggled to gain market traction and replace the traditional walking cane as the standard assistive tool for people with visual impairments.
Current technological limitations include prohibitive costs, bulky hardware requirements and non-user-friendly interfaces~\cite{golledge2004stated,yusif2016older,arditi2013user}.
To address these issues, we implemented a handheld mobile system that is based on a concept proposed in~\cite{lock2017portable,lock2019active} and tested in~\cite{lock2019bone} that uses a Google Tango device that is able to localise itself in real-time.
This system has the benefit of minimal hardware requirements and a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.

@@

\subsection{Audio Interface}

@@

@@

\subsubsection{Pan Dimension}

@@

\subsubsection{Elevation Dimension}

@@

\section{Experiments}

@@

\subsection{Participant Characterisation}

In addition to the target searching experiment, another set of experiments were conducted to characterise the participants' hearing characteristics.
The characteristics that were measured were each participant's audio localisation ability on the lateral plane, as well as the participants' ability to discriminate between tones with different frequencies. 
These results will be used to provide context to the results captured during the target search experiment and provide additional insight on any possible biases or limitations.

\subsubsection{Sound Localisation}

In this experiment, we evaluated a participant's ability to determine the lateral direction a sound is coming from.
To do this, we played a continuous \SI{512}{\hertz} sinusoidal tone to the participant through the headphones and applied an HRTF to spatialise and place its source to the participant's left or right.
The participant then had to select the direction the sound came from.
The longer the experiment lasts and the more correct guesses the participant makes, the closer the source moves to the centre-front of the participant, making it progressively harder to localise.

For this progressive increase in difficulty, a 2-up, 1-down step process is used~\cite{wetherill1965sequential,levitt1971transformed}, meaning that for every 2 correct answers, the distance to the centre halves.
Conversely, the task becomes easier for each incorrect answer by doubling the sound source's distance from the centre.
We also use 2 different step sequences, one starting at a large distance (\SI{45}{\degree}) from the user and the other at the minimum distance (approximately \SI{1}{\degree}), giving an `easy' and a `hard' progression respectively.
The terminating condition for the experiment is when the 2 sequences converge to within 2 intervals of one another for 3 consecutive guesses.
This gives a distance band within which the participant is capable of localising the sound source.
Each participant performed this experiment three times.

\subsubsection{Pitch Discrimination}

Here we determined a participant's ability to differentiate the pitches of 2 tones being played to them, i.e.\ how well they can tell if a tone is high or low pitched.
We played 2 tones to the participants in succession with the second tone being higher or lower-pitched than the first.
The participants were then asked to select whether they interpreted the second tone as higher or lower-pitched than the first.

The first tone is randomly generated, while the second tone is generated by adding or subtracting some value from the first one.
The tone difference depends on how well the participant can tell the tones apart.
Like the sound localisation experiment, a 2-up, 1-down step process is used: for every 2 consecutive correct answers, the pitch difference between the tones is halved and is doubled for every incorrect answer.
2 step sequences are again used here, one starting with a large pitch difference ($f_h=2^9=$ \SI{512}{\hertz}) between the tones and the other with a small difference ($f_l=2^1=$ \SI{2}{\hertz}).
The termination condition is when the 2 step sequences are within 1 octave of each other (i.e.\ $\log_2\frac{f_h}{f_l}=2^1$) for 3 consecutive answers.
Each participant performed this experiment twice.

\subsection{Target Search}

To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
The participants were given a Tango device running an app written for the experiment that generates a set of virtual targets and presents them to each participants, one at a time. 
The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
When a participant was confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localisation API.\
A total of 28 targets were generated per participant.

\subsection{Metrics}

We use two different metrics to compare the three different pitch gradient settings: the acquisition accuracy and search time.
The accuracy is given as the difference between the Tango's orientation at the time the participant confirmed they were on target, and the target's actual orientation.
We separate the results of the tilt and pan dimensions in order to see how the different pitch gradients affect a participant's pointing accuracy.

We also compare the performance of the three pitch gradient settings in terms of the time it takes each participant to find a target.
However, since each participant was presented with a different, randomly generated set of targets, a direct time comparison is not possible.
Instead, we use Fitts's Law~\cite{fitts1954information}, modified by MacKenzie for uncertain target sizes and noisy data~\cite{mackenzie1992fitts}, which states that there is a relation between the time it takes to find a target and its index of difficulty (the ratio between the distance to the target and its width).
It also provides a so-called `index of performance' that we can use as a metric to compare the results between the three configurations.

Here we briefly summarise the equations and quantities involved in our metric.
Fitts's Law is given by

\begin{equation}
  \label{eq:fitts-base}
  t = a + bID,
\end{equation}

\noindent
where $t$ is the time it takes to find a target, $a$ and $b$ are constants determined through regression and $ID$ is a description of the difficulty of the target, given as logarithmic function of the ratio between the distance to the target and the target's width.
In our case, the targets have no width, since they are points in space, and we therefore use MacKenzie's modified form for $ID$, given by

\begin{equation}
  \label{eq:fitts-id}
  ID = \log_2\left(\frac{\theta}{w_e} + 1\right).
\end{equation}

\noindent
Here $\theta$ is the angular distance between subsequent target centres and $w_e$ is the targets' effective angular width~\cite{welford1968fundamentals}, given by

\begin{equation}
  \label{eq:fitts-we}
  w_e = \sqrt{2\pi e}\sigma = 4.133\sigma,
\end{equation}

\noindent
where $\sigma$ is the standard deviation of the error data, taken as the angle between the participant's target selection and target's actual angular position.
Fitts's index of performance, $IP$, can then be calculated using

\begin{equation}
  \label{eq:fitts-performance}
  IP = \frac{ID}{t},
\end{equation}

\noindent
where $IP = \frac{1}{t}$ when $\sigma=0$.

\subsection{Procedure}

Two groups of participants were recruited for the experiments. 
Group \textit{G1} consisted of 42 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 12 people with severe visual impairments. 
Both groups were given some time before the experiments to familiarise with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets, without worrying about the time it took.

\section{Results}

\section{Conclusion}
